Landcover Data Limits Our Understanding of Global Change

### 2015-11-02 paper/main.tex
2/11/2015 commit. See changelog

\title{Landcover Data Limits Our Understanding of Earth System Processes}
%\title{Our Understanding of Global Change is Built on a Shaky Foundation}

@@

\begin{article}
\begin{abstract}
{Blah blah.}
\end{abstract}

@@

\dropcap{T}he functioning of the Earth System\footnote{give definition here} is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface. Human endeavors are both strongly governed by and shape landcover, whether it be felling ancient forests for timber production or burning savannas to stimulate a green flush for livestock, while landcover is a primary driver of climate and biogeochemical processes \cite{lambin_dynamics_2003}. The vastness of our modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover. To fully understand these processes and the nature of global change, it is therefore essential to know the nature and distribution of landcover.

This importance is understood by a growing number of social, economic, and physical scientists, who increasingly use landcover data to advance understanding in Earth System research areas ranging from food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010} to carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, and demographic shifts \cite{linard_assessing_2010}.

The validity of the knowledge resulting from such studies depends on the veracity of the landcover data upon which they are based, much as a building requires a solid foundation if it is to remain standing.  Unfortunately, the evidence so far suggests that Earth System science is being built on a shaky foundation. The reason for this is that landcover data can only practically be derived from satellite imagery, but in many regions the cover types of interest are smaller \cite[e.g. smallholder's farms][]{jain_mapping_2013} than the sensor resolution, or spectrally indistinct from neighboring covers, which translates into substantial mapping errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}. The result is that landcover maps are generally inaccurate at finer scales and disagree substantially with one another, particularly in those parts of the world undergoing the most rapid land use changes \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}. These errors mean that we are still unable to obtain the granular, mechanistic understanding of global change processes that we need.

These problems with landcover products are known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}. What remains an open question is exactly how much the maps Earth System researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of Earth System processes. Answering this question depends on having spatially comprehensive ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}. Our understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data. This allows us to quantify between map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.

@@

In this study, we take advantage of these recent advances to address the urgent need to more thoroughly and precisely quantify landcover map errors and how they might impact our understanding of Earth System processes in the world's most dynamic regions.  We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through ``downstream'' studies investigating into both the physical and socioeconomic components of the Earth System. Our objective is to provide scientists and policy-makers who use landcover data with a better, more up-to-date, understanding of their appropriate uses and limitations.

@@

\vspace{-0.75 cm}
\begin{figure}[h]
\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model. Prediction curves are color-coded to the different test maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}\label{afoto2}
\end{figure}

@@

@@

\vspace{-0.75 cm}
\begin{figure}[ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scalew.pdf}}
\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map. Bias estimates (represented by symbols) fall within the semi-transparent floating bars, while accuracies are contained in the solid bars. Bar colors are coded to specific cropland map, symbols indicate which cover type was used to calculate cropland-adjacent carbon density. The bar represents the mean biases calculated across each of the 5 cover types. Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2). MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.}
\label{afoto}
\end{figure}

\subsection{Evapotranspiration estimates}
Accurate estimation of hydrological fluxes is critical to understanding how land-atmosphere interactions impact the climate system and runoff \cite{liang_simple_1994}. Land surface hydrological models, such as the Variable Infiltration Capacity \cite{liang_simple_1994}, are used to simulate these processes, and depend on landcover maps to provide information on the characteristics of vegetation and other materials covering the surface, as these govern the rates of runoff, infiltration, and evapo-transpiration. We used the VIC model to generate 25 km estimates of monthly evapotranspiration throughout South Africa, and examined how these were impacted by error in the test maps, which were used to determine the landcover-specific leaf area index (LAI) values that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.

@@

\vspace{-0.75 cm}
\begin{figure}[h]
\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
\end{figure}

@@

%\vspace{-0.5 cm}
\begin{figure}[!h]
\vspace{-0.75 cm}
\centerline{\includegraphics[width=0.5\textwidth]{figures/yield_prod_bias.pdf}}
\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates. Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.  Symbols code the different variables (production and yield), normalized to their respective means.}
\label{afoto}
\end{figure}

\subsection{Initializing an agent-based model}
Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}. A common feature of such ABMs is that they need to be calibrated against data describing the characteristics of land users, including an initialization step to assign land resources to ``agents'' representing the land users, wherein the simulated landscape pattern and distribution of agent resources matches those in the real world. In our example, we used an ABM of household food security that simulate the interactions between many individual farming households (the agents) and their environment over multiple seasons \cite{chen_dependency_2013}. We used cropland maps to provide the model the location and abundance of cropland, which is used to allocate an initial share of cropland to each simulated household. Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g. districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers. To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig. S8) in the eastern part of the country with 28-45\% of their areas devoted to cropland. The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity. The number of households and the cropland area per household is derived from survey data of communities where all cropland is owned. The model is thus considered adequately initialized when all households are allocated their appropriate area of cropland, and all cropland is occupied.

We used the reference map and each cropland map to separately initialize the model, and compared the agent allocation results to assess how cropland map errors impacted the initialization process. We examined three variable, the first being the number of agents that were not assigned fields, the second the amount of cropland left unallocated, and the third the area of land deficit, or the amount of land that should have been assigned to households but wasn't. For the first variable, there was a one-to-one relationship between the percentages by which cropland was underestimated and households that could not be assigned fields (Fig. 6, left panel). The most extreme examples occurred when MODIS cropland initialized the ABM in districts 1 and 2, where $\sim$85\% of agents did not receive cropland. All households were assigned fields when total cropland area was overestimated (GeoWiki, SA-LC), but in these cases the area of cropland allocated to no one (the second variable) was proportional to the size of the overestimate (e.g. $\sim$20\% for SA-LC, Fig. 6 right panel). Interestingly, the overall relationship between the percent of cropland allocated and percent cropland error was U-shaped, as the model also failed to give land to households when cropland was underestimated by more than 50\% (Fig. 6, right panel). MODIS again provided the most pronounced results in districts 1 and 2, where 7-12\% of cropland was left unallocated despite the fact that 85\% of agents had no land. This curious relationship occurred because cropland tends to cluster, and when it is underestimated, the size of these clusters is small, resulting in islands of cropland that fall outside of the search radius (which is constrained by an absolute distance and the proximity of other agents) within which cropland is sought when agents are seeded onto the landscape.

The last measure, land deficit, increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2--and would become infinite in the case of a 100\% underestimate.

@@

\begin{figure}[!ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/agent-bias.pdf}}
\caption{Biases in agent-based model initialization relative to the district-wise errors (as a percent) in total cropland area, measured in terms of the percent of households having no cropland allocated (left), and the percent of cropland left unallocated (right). Dot sizes correspond to district numbers, colors represent the landcover map.}
\label{afoto}
\end{figure}

\section{Discussion}
This spatially comprehensive, bottom-up assessment of landcover map bias and inaccuracy and provides unique insight into their extent, causes, and consequences for understanding global change processes, made possible by a unique, high accuracy dataset that likely provides the truest measure of total cropland area and distribution that is currently available for this region. This dataset is of course not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission), while some of the cropland map error we found may have been caused by the slight temporal mismatches between the reference data and the original landcover datasets we used. However, our assessment (SI) suggests that these errors are small, and do not appreciably impact our findings, which is bolstered by previous work showing the large scale of disagreements between landcover map-based cropland area estimates and national inventory data \cite{fritz_comparison_2010}.

Our results suggest several guidelines for using landcover data in global change research, and contain some important implications for how understanding of global change processes based on the data, and associated policy decisions, may be affected. In terms of developing a base landcover map, the first rule of thumb is that standard landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, appear to be too biased to be useful without substantial aggregation. If we use the standard that bias within +/-10\% is acceptable, then at least 25-100 km of aggregation is needed to sufficiently cancel out the errors in the base landcover data and subsequent first order estimates built on them (Fig. 3 \& 4). The upper range of aggregation scale is necessary if a mixed pixel class becomes dominant, as in the case with GlobCover, because these lead to underestimation bias that will persist until the pixel size becomes substantially greater than the average area of landscapes that are dominated by the cover type of interest, which can be $>$1000 km$^2$ in some of South Africa's farming regions.

Maps derived from higher resolution sensors, such as the SA-LC dataset, if carefully done, do not have this mixed class problem, and are sufficiently unbiased for most applications with just 1-5 km of aggregation. However, such datasets are typically developed for specific countries, using varying methods, and can be hard to obtain. For broader scale analyses, the best option is to use newer generation maps such as GeoWiki (and the GLC-Share \footnote{GLC-Share. www.glcn.org} datasets for other cover types) which is relatively unbiased at 1 km resolution. GeoWiki's lower bias comes from its process of evaluating consensus between several landcover datasets (including the other three in this study), resulting in cropland probabilities that are converted to percentages by calibrating to statistical data \cite{fritz_cropland_2011,fritz_mapping_2015}. This method mirrors the ensemble averaging used by other fields (e.g. crop \cite{asseng_uncertainty_2013}, climate \cite{giorgi_calculation_2002}, and ecological modeling \cite{araujo_ensemble_2007}) to increase prediction confidence.

GeoWiki's statistical constraint procedure is similar to the one we used \cite[following][]{ramankutty_farming_2008}, which produced unbiased maize production estimates (Fig. 3) by eliminating bias in the adjusted cropland and harvested area maps that they were built upon. This result, together with GeoWiki's low bias, indicate the value of fusing inventory data with remote sensing. However, this method depends on the quality of inventory data, which are often suspect, particularly in Africa \cite{carletto_emperor_2013,fao_action_2013}. The statistical constraint also does not greatly improve map \emph{accuracy}, as evidenced by GeoWiki's 23\% mean absolute error in 1 km cropland percentage estimates (Fig. S1), which is only slightly more accurate than MODIS (31\%) but worse than SA-LC (11\%). GeoWiki is definitely most accurate among the large scale landcover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier. Statistical constraints only adjust map values at locations where cropland is identified, so their use it .

Map accuracy is perhaps more important than bias for landcover maps.

Broader regional implications - error higher elsewhere

Main points:

What we found, significance of study 
\begin{itemize}
  \item First large area quantification of spatial biases
  \item How large those biases are, for one of the most widely spread (spreading landcovers)
  \item Insight into causes of bias, and thus some understanding of where biases are likely to be greater or smaller
  \item How much progress made in reducing it
  \item Class type and bias
\end{itemize}

\begin{itemize}
  \item Bias decreases as function of scale
  \item General bias patterns, appropriate use of landcover products, which landcover products
  \item Appropriate scales of inference, by type of product - 
  \item Aggregation improves results for landcover, generally, 
  \item Sensor resolution, statistical resolution, and merging products have high value
   \begin{itemize}
     \item But don't remove spatial bias - absolute bias matters. Statistical constraints seems to just compress spatial biases to higher rates of turnover. Geo-wiki
     \item But of course these types of data are then dependent on how accurate the statistical data are defining the constraint (cite emperor has no data)
    \end{itemize}
   \item Mixed landscapes increase the chances of omission and commission errors by increasing the number of cover classes, or because such landscapes are less spectrally distinct \cite{estes_diylandcover:_2015}
   \item caveats: Only single country in South Africa. More commercial farming than many other countries, but results are still instructive. Analysis of error as function of landscape type suggests that areas where cropland is more mixed with natural vegetation have higher errors. These sorts of landscapes quite common in smallholder-dominated systems, thus suggests that biases may be even higher elsewhere on the continent. 
\end{itemize}

Implications for understanding global change and policy: \\

Increasing awareness of need to have spatial assessments in global change analyses. Do things such as identify areas where yield gaps are high, or how much carbon or biodiversity will be lost to changes in land use, in order to try prioritize development \cite{searchinger_high_2015,newbold_global_2015}, or to understand coupled human-social dynamics, etc. \\

Our finding suggest:

\begin{itemize}
  \item Area-based estimates only safe at coarser scales of aggregation for most types of global change analyses, and primarily with constrained products. 
  \item 50-100 km scale of aggregation reduces bias sufficiently.
  \item Not so with unconstrained products
  \item Assessments of spatial variability unsafe, for all products, bar one - finer country-scale product.  Here you look at absolute bias. This is high in many products even at higher scales of aggregation. 
  \item This suggests that disaggregation approaches or paint by numbers approaches are nice maps, but can't give clear guidance about differences between grid cells, even when highly aggregated.  [work on this] 
  \item can lead to misinformed policies
  \begin{itemize}
    \item E.g. Efforts to identify area where yield gaps are most pronounced and/or concentrated are likely to be highly misleading, leading to ineffective targeting of resources. Most informative simply to look at these areas at the political boundary resolution
    \item Comparing carbon stocks against potential yield for tradeoff analysis, which may be done with conservation planning to find areas with high benefit/low-cost. Also misleading. 
    \item Looking at land availability for cropland or biofuels expansion (look at biofuel paper for example)--land might not be as available as people think. Can lead to formulation of bad policy
    \end{itemize}
  \item Analyses of higher order interactions, biogeochemistry, human decision-making, also misleading (maybe pair this with yield example). 
  \begin{itemize}
     \item Our example here, ET estimates not heavily biased, but in marginal areas of low rainfall some pronounced differences. These are areas where irrigation is more common, but VIC doesn't simulate this, so absolute bias in those zones likely to be underestimated, and such regions can have substantial impacts on altering climate \cite{estes_changing_2014, sacks_effects_2008}. 
    \item Can skew understanding of more advanced attempts to understand the human factors that go into driving agricultural productivity. Examples here
    \end{itemize}
\end{itemize}

Way forward
\begin{itemize}
  \item For now, use latest generation products fusion products or more detailed country-level products
  \item Avoid change detection based on landcover products, e.g. MODIS. 
  \item But moving forward key will be developing new approaches to map landcover with much greater fidelity, e.g. scaling out approach that led to this dataset, combining with latest computer vision algorithms, etc.  
\end{itemize}

>>>>

\title{Landcover Data Limits Our Understanding of Global Change}
%\title{Our Understanding of Global Change is Built on a Shaky Foundation}

@@

\begin{article}
\begin{abstract}
{Guidelines: landcover maps based on coarser resolution individual sensors (e.g. MODIS) should be aggregated to 50-100 km resolution to minimize map bias and accuracy. Map accuracy is lowest at intermediate levels of cover, thus in highly mixed landscapes, such as small-holder farming systems or ecotones, even coarser aggregation may be necessary. Finer resolutions can be used with Landsat-based maps, if available, or for global analysis from new improved fusion products, particularly GeoWiki. The widely used datasets for carbon, crop area, and crop yield should be rebuilt on these updated datasets. Newer approaches to improving landcover maps, using interactive training with expert judgement and next generation classifiers are needed to create more reliable landcover, and in doing so will improve our understanding of global change. }
\end{abstract}

@@

\dropcap{T}he functioning of the Earth System is fundamentally connected to the characteristics of landcover \cite{lambin_modelling_1997}, the physical constituents of the terrestrial surface. Human endeavors are both strongly governed by and shape landcover, whether it be felling ancient forests for timber production or burning savannas to stimulate a green flush for livestock, while landcover is a primary driver of climate and biogeochemical processes \cite{lambin_dynamics_2003}. The vastness of our modification of the Earth's surface \cite{lambin_dynamics_2003} means that socioeconomic and physical processes increasingly interact through landcover. To fully understand these processes and the nature of global change, it is therefore essential to know the nature and distribution of landcover.

This importance is understood by a growing number of social, economic, and physical scientists, who increasingly use landcover data to advance understanding of global change in areas ranging from food security \cite{lark_cropland_2015,wright_recent_2013, licker_mind_2010} to carbon cycling \cite{asner_high-resolution_2010, gaveau_major_2014}, biodiversity loss \cite{newbold_global_2015, luoto_predicting_2004}, and demographic shifts \cite{linard_assessing_2010}.

The value of the insights resulting from such studies depends on the veracity of the landcover data upon which they are based, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations. The reason for this is that landcover data can only practically be derived from satellite imagery, but in many regions the cover types of interest are smaller \cite[e.g. smallholder's farms][]{jain_mapping_2013} than the sensor resolution, or spectrally indistinct from neighboring covers, which translates into substantial mapping errors \cite{see_improved_2015,lobell_use_2013,estes_diylandcover:_2015}. The result is that landcover maps are generally inaccurate at finer scales and disagree substantially with one another, particularly in those parts of the world undergoing the most rapid land use changes \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}. These errors mean that we are still unable to obtain the granular, mechanistic understanding of global change processes that we need.

These problems with landcover products are known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway, particularly for agriculture \cite{fritz_geo-wiki:_2012,estes_diylandcover:_2015}. What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes. Answering this question depends on having spatially comprehensive ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}. Our understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data. This allows us to quantify between map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.

@@

In this study, we take advantage of these recent advances to address the urgent need to more thoroughly and precisely quantify landcover map errors and how they might impact our understanding of change in the world's most dynamic regions.  We use a unique, high accuracy landcover map of South African crop fields to conduct a spatially comprehensive, bottom-up quantification of error in several widely used landcover maps, and how these errors can propagate through ``downstream'' studies investigating into both the physical and socioeconomic components of global change. Our objective is to provide scientists and policy-makers who use landcover data with a better, more up-to-date, understanding of their appropriate uses and limitations.

@@

@@

%\vspace{-0.75 cm}
\begin{figure}[!h]
\centerline{\includegraphics[width=.5\textwidth]{figures/biases_md_lnorm_gam_mu0.pdf}}
\caption{The relationship between map accuracy (the mean absolute error) in test maps and the actual cropland cover within agricultural landscapes (reference map pixels having $>$0.5\% cropland), here defined by the boundaries of magisterial districts (n = 345), as fit with a generalized additive model. Prediction curves are color-coded to the different test maps, with the solid line indicating predicted absolute bias, and the lighter shading the standard error of the coefficients.}\label{afoto2}
\end{figure}

@@

%\vspace{-0.75 cm}
\begin{figure}[!h]
\centerline{\includegraphics[width=.5\textwidth]{figures/carbon_veg_scalew.pdf}}
\caption{Biases and accuracies (mean absolute errors) of carbon densities derived from cropland maps, calculated as percents relative to the reference map. Bias estimates (represented by symbols) fall within the semi-transparent floating bars, while accuracies are contained in the solid bars. Bar colors are coded to specific cropland map, symbols indicate which cover type was used to calculate cropland-adjacent carbon density. The bar represents the mean biases calculated across each of the 5 cover types. Shrubland and grassland bias values were near zero, while secondary forest values were close to forest values, and thus these are not shown for display clarity (but see Table S2). MODIS and GlobCover values at 1 km exceeding the plot's Y limits are provided near their truncated tops.}
\label{afoto}
\end{figure}

\subsection{Evapotranspiration estimates}
Accurate estimation of hydrological fluxes is critical to understanding how land-atmosphere interactions impact the climate system and runoff \cite{liang_simple_1994}. Land surface hydrological models, such as the Variable Infiltration Capacity \cite{liang_simple_1994}, are used to simulate these processes, and depend on landcover maps to provide information on the characteristics of vegetation and other materials covering the surface, as these govern the rates of runoff, infiltration, and evapo-transpiration. We used the VIC model to generate 25 km estimates of monthly evapotranspiration throughout South Africa, and examined how these were impacted by error in the test maps, which were used to determine landcover-specific values for leaf area index (LAI), plant rooting depth, aerodynamic roughness, and several other variables that VIC uses to partition water vapor fluxes into their evaporative and transpirative components.

@@

%\vspace{-0.75 cm}
\begin{figure}[!h]
\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
\end{figure}

@@

\subsection{Agent-based simulation of food security}
Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}. To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistic represents the social and biophysical features of the study region \cite{berger_creating_2006}. In our example, we used an ABM of household food security that simulates the food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}. The model is initialized such that each household is allocated its designated share of cropland, and then produces a crop over the course of one to many seasons. Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g. districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers. To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, Fig. S8) in the eastern part of the country with 28-45\% of their areas devoted to cropland. The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity. The number of households and the cropland area per household is derived from survey data of communities where all cropland is owned. The model is thus considered to well calibrated when all households are allocated their appropriate area of cropland, and all cropland is occupied.

%\vspace{-0.5 cm}
\begin{figure}[!hb]
%\vspace{-0.75 cm}
\centerline{\includegraphics[width=0.5\textwidth]{figures/yield_prod_bias.pdf}}
\caption{Bias (mean error) and accuracy (mean absolute error [MAE]) in disaggregated maize yield and production estimates. Bias estimates (represented by symbols) fall within the semi-transparent bars, mean absolute errors in the solid bars, with bar colors coded to specific cropland maps.  Symbols code the different variables (production and yield), normalized to their respective means.}
\label{afoto}
\end{figure}

We used the reference map and each cropland map to separately initialize the model, and examined how map errors impacted the land allocation process and household food production estimates. We examined three variables: the percent of unallocated cropland, land deficit, and food deficit. The percent of unallocated cropland measures how effective the model was in matching household agents to available cropland resources. Where error was negative (indicating a cropland overestimate by the test maps), the amount of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig. 6A). When cropland was underestimated, no land was left unallocated until the underestimation exceeded 50\%. The MODIS-based simulation for districts 1 and 2 was most pronounced for this tendency, where 5-10\% of cropland remaining unallocated despite the fact that cropland with such a high cropland underestimate (85\%) the majority of households were not allocated cropland. This curious relationship occurred because croplands tend to cluster, and when underestimated clusters tend to be small, resulting in islands of cropland that fall outside of the search radius within which cropland is sought when agents are seeded onto the landscape.

The second measure, land deficit, describes the amount of land that the model should have assigned but couldn't due to the mismatch between the cropland map and the statistical estimate of cropland holdings. Land deficit increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig. 6B)--and would become infinite in the case of a 100\% underestimate. This contrasted with the third measure, food deficit, the shortfall in the average amount of food that should have been produced by each household but wasn't because of the land deficit; this particular deficit increased linearly with the percentage of cropland underestimate (Fig. 6C).

@@

\begin{figure}[!ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/figure6.pdf}}
\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the percent of cropland that should have been allocated to households but wasn't, and C) the food deficit, or the average food production across households resulting from inadequate cropland allocation. Dot sizes correspond to district numbers, colors represent the landcover map.}
\label{afoto}
\end{figure}

\section{Discussion}
This spatially comprehensive, bottom-up assessment of landcover map bias and accuracy provides new insight into their extent, causes, and consequences for understanding global change, made possible by a unique, high accuracy dataset that likely provides the truest measure of cropland area and distribution currently available for this region. This dataset is not perfect, being affected by the map-makers' occasional interpretation errors (mostly of omission), while some of the error we identified may be due to slight temporal mismatches between the reference and test datasets. However, our assessment (SI) suggests that these errors are small, and do not appreciably impact our findings, which is bolstered by previous work showing the substantial errors and inconsistencies between landcover maps \cite{fritz_comparison_2010,gross_monitoring_2013}. These previous studies, which were conducted over broader extents than ours, also indicate that our findings should be generalizable beyond South Africa's borders, particularly to more rapidly developing, data poor regions.

\subsection{General guidelines for selecting and using landcover data}
Our results suggest several guidelines for selecting and using landcover data in global change research. First, landcover products derived from coarse resolution sensors, such as MODIS and GlobCover, need substantial aggregation to reduce the large biases and inaccuracies that manifest at finer resolutions. Researchers might need to use these datasets for a variety of reasons, for instance, because they provide consistent, multiple landcover classes at a global scale, or allow changes to be tracked across large scales over time \cite{luoto_predicting_2004}. The appropriate scale of aggregation then depends on which error property the user wants to minimize. For instance, if the objective is to minimize systematic errors in order to calculate an unbiased mean (e.g. average carbon density), then at least 50-100 km of aggregation is needed to reduce map bias to within +/-10\% (Fig. 1B, 3). If the user wants to maximize accuracy (accuracy is important for comparing differences between locations on a single map, or between dates at fixed locations), then even coarser aggregations are needed to reduce mean absolute error below 10\% (Fig. 1, 3, 5).

The characteristics of cover at the landscape scale may also require greater levels of aggregation to achieve sufficient accuracy. We found that accuracy is lowest where mixing of cover types is greatest (Fig. 2), a result echoed in an Africa-wide change detection study made with MODIS \cite{gross_monitoring_2013}. Outside of South Africa, where smallholder-driven agriculture predominates \cite{lambin_estimating_????} such mixed landscapes are more common, thus greater levels of aggregation may be needed over these areas. Coarser aggregations are also necessary if the cover type of interest belongs primarily to a mixed class (as with GlobCover). This leads to underestimation bias (Fig. 2) and high inaccuracy, which will persist until the aggregated pixel exceeds the characteristic size of the geographical area within which the cover type dominates. In South Africa's farming regions, this can exceed 1000 km$^2$.

Maps derived from higher resolution sensors, such as the SA-LC dataset, do not have this mixture problem because the pixels are fine enough to be assigned to a single cover type. This characteristic leads to higher accuracy, which is what we observed here: the SA-LC dataset needed the least aggregation--just 5-10 km--to achieve $<$10\% MAE in most of the example applications we tested (note, however, another guideline in this result: even with high resolution maps some aggregation is needed to achieve higher accuracy).

%needed This observation leads to a second rule of thumb: users should select datasets from higher resolution sensors where possible (assuming they were carefully developed), as these needed much less aggregation to return accurate estimates for most applications we tested here. For $<$10\% MAE, just 5-10 km of aggregation were needed (however, this is still much coarser than the 30 m resolution of the sensor itself).

Unfortunately, Landsat-scaled maps are typically developed for specific countries, using varying methods, and can be hard to obtain (an exception to this is the newly released GLC30, which has a reported classification accuracy of 80\% \cite{chen_global_2015}). Our results also suggest that, despite their higher accuracy, they are not necessarily the least biased maps. Except for a few results (notably the 1 km carbon and maize production estimates), GeoWiki-based maps had the least amount of bias across most scales. GeoWiki's lower bias results from its process of calibrating the cropland percentages to reported agricultural statistics \cite{fritz_cropland_2011,fritz_mapping_2015}.  The procedure is similar to the one we used in disaggregating maize harvested areas \cite{ramankutty_farming_2008, monfreda_farming_2008}, which led to fairly unbiased estimates of maize production and yield (Fig. 5) above 1 km (except for GlobCover's production estimates). This result, together with GeoWiki's unbiased cropland estimates (Fig 1.B), indicates the value of fusing inventory data with remote sensing to reduce map bias.

%The consensus method mirrors the ensemble averaging used by other fields (e.g. crop \cite{asseng_uncertainty_2013}, climate \cite{giorgi_calculation_2002}, and ecological modeling \cite{araujo_ensemble_2007}) to increase prediction confidence.

Of course, for the constraint to be effective, the inventory data themselves must be faithful to reality, but in many places, particularly Africa, agricultural statistics are suspect \cite{carletto_emperor_2013, fao_action_2013}, which is a vulnerability of the fusion approach \cite{see_improved_2015}. Beyond this limitation, this method does not improve map accuracy, which is evident in GeoWiki's relatively high MAE for cropland percentage at 1 km (23\%, Fig. 1B), and by all maps' (except SA-LC's) highly inaccurate maize production estimates (Fig. 5). The reason that accuracy is not improved is because the statistical constraint cannot correct the omission and commission errors within the landcover maps.

These last observations suggest two further guidelines: 1) for high accuracy, users should select well-validated datasets from higher resolution sensors--this option should be increasingly available as computational power grows and high resolution image archives expand \cite{hansen_high-resolution_2013,chen_global_2015}; 2) where low bias is needed over large areas, the best option is to use newer generation, statistically constrained maps such as GeoWiki (and possibly the GLC-Share \footnote{GLC-Share. www.glcn.org} datasets for other cover types). But in either case some aggregation--5-25 km, depending on application and whether low bias or high accuracy is the objective--will still likely be necessary to achieve sufficiently low error. Additionally, these higher accuracy, lower bias product are made infrequently, thus applications requiring high frequency landcover data will depend on MODIS and confreres, as previously mentioned, and therefore much coarser aggregations.

%GeoWiki was the most accurate among the large scale landcover products, but its improvement is related to the map consensus methods, which can correct for omission or commission errors made by the classifier.

\subsection{Implications for global change analyses}
Besides the preceding landcover data selection and usage guidelines, our findings also have broader implications for global change analyses and associated policy decisions. The simpler, most direct uses of landcover data have substantial to skew our understanding of global change processes. For example, the global carbon carbon density map of \cite{ruesch_new_2008}, of which we created facsimiles here, has been widely used, but our results suggest that any analyses of the spatial variability of carbon stocks, or of mean carbon densities, can be highly misleading if the map is not first sufficiently aggregated. For example, a study of deforestation risks in the Democratic Republic of the Congo estimated forest rents as a function of carbon densities based on this map. Although the map was aggregated to $\sim$50 km, we found that mean absolute errors can still reach 40\% at this scale (Fig. 3). The same carbon map provided country-specific mean forest carbon densities for an evaluation of climate mitigation policies \cite{cattaneo_international_2010}, which our results suggest could be underestimated by 70\% (Fig. 3). In contrast to these, an assessment of the congruence between carbon and biodiversity \cite{strassburg_global_2010} aggregated the carbon data to $\sim$110 km resolution, where accuracy is much higher.

A related example can be seen in studies that create high resolution maps (10 km) showing the potential food production benefit of closing yield gaps \cite[e.g. Figure 3 in][]{foley_solutions_2011}. Such maps give an inflated sense of the precision with which such locations can be identified. Although the study in this example is careful to report aggregated statistics, in which we can have greater confidence, it is not inconceivable that the maps themselves are used to inform policy. For this reason, it is advisable to degrade maps until their accuracy is of an acceptable standard (i.e. $\geq$100 km).
